\bf Nonparametric Block Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test

### 2025-02-18 manuscript/manuscript.tex
politis2004blksize

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.  
Block-bootstrap can be done with overlapping or moving blocks.
Define moving blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study. \citet{hall1995blocking} and 
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. However,
because this method is intended for cases where the dependence structure is 
unknown, the block size choice will not vary with the level of dependence.
Now we draw $k$ blocks from the $(n - l + 1)$ blocks 
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block 
selected will be reduced in size so that the final size of the block bootstrap 
sample is $n$.

@@

In summary, the procedure of the NPBB test is 
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying moving block bootstrap 
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters 
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where 
  $C_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] - 
  F(x; \theta_0^*))$ is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

Several promising directions for further research emerge from the findings of
this study. Firstly, there are many other variants of bootstrap besides
moving block bootstrap which
can be used for serially dependent data such as the circular block boostrap \citep{romano1992circular}, the 
stationary bootstrap \citep{politis1994stationary}, and the dependent wild 
bootstrap \citep{shao2010dependent}. It would be interesting to investigate how
these methods perform using our correction. Secondly, analogous 
methodologies for other goodness-of-fit tests
based on empirical distributions, such as the Anderson--Darling and Cramér--von
Mises tests---which have been shown to outperform the KS test in certain scenarios
\citep{stephens2017tests}---could be explored. The application of block
bootstrapping bias correction to these tests could be particularly beneficial in
environmental studies, such as extreme value modeling, where observations are
often serially dependent. Thirdly, an extension of the proposed NPBB
one-sample KS test to a two-sample setting would be
intriguing. This adaptation would test whether two samples originate from the
same distribution without requiring specification of serial dependencies in
either sample, maintaining the advantageous feature of the original method.
Finally, a rigorous theoretical justification of the NPBB method for
stationary series is of significant academic interest. Establishing
precise conditions for the asymptotic behavior of the KS statistic and
the consistency of the bootstrap procedure would offer valuable
practical insights into the scenarios where the NPBB approach is
expected to perform optimally.

>>>>

We now consider the case where $X_i$'s are realizations from a time series and
$X^{(b)}_1,...,X^{(b)}_n$ are generated by block bootstrap for 
$b \in \{1, \ldots, B\}$.  
Circular block-bootstrap can be done with overlapping blocks that can
wrap around the end of the series \citet{romano1992circular}.
Define blocks (assuming block size $l > 1$) as:
\begin{equation*}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation*}
A common 
function for block size that is considered optimal is 
$l = \lceil n^{1/3} \rceil$ \citep{buhlmann1999block},  
which was adopted in this study. \citet{hall1995blocking} and 
\citet{politis2004automatic} note that
the selection of block size should ideally take into account the strength
and direction of serial dependence. However,
because this method is intended for cases where the dependence structure is 
unknown, the block size choice will not vary with the level of dependence.
Now we draw $k$ blocks from the $(n - l + 1)$ blocks 
of $Z_j$'s with replacement and then align them in the order they were picked to
form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last block 
selected will be reduced in size so that the final size of the block bootstrap 
sample is $n$.

@@

In summary, the procedure of the NPBB test is 
summarized as follows. Repeat the following steps for $b \in \{1, ..., B\}$.
\begin{enumerate}
\item
  Generate $X^{(b)}_1,...,X^{(b)}_n$ by applying circular block bootstrap 
  on the original sample as
  defined previously.
\item
  Obtain parametrically fitted parameters 
  $\hat\theta^{(b)}_n$ of $\theta$ from $X^{(b)}_1,...,X^{(b)}_n$.  
\item
  Obtain the empirical distribution function $F^{(b)}_n$ of
  $X^{(b)}_1,...,X^{(b)}_n$. 
\item
  Calculate bootstrap KS statistic
  \[
    T^{(b)}_n = \sup_x \Big\vert \sqrt{n}\left(F^{(b)}_n(x) 
    - F(x; \hat\theta^{(b)}_n)\right) - C_n(x) \Big\vert.
  \]
  where 
  $C_{n}(x) = \sqrt{n}(\mathbb{E}^{*}[F^{(b)}_n(x)] - 
  F(x; \theta_0^*))$ is the bias term, and
  the expected values $\mathbb{E}^{*}[F^{(b)}_n(x)]$ and
$\theta_0^* = \mathbb{E}^{*}[\hat\theta^{(b)}_n]$ can be
approximated by, respectively,
$\mathbb{E}_B^{*}[F^{(b)}_n(x)] = \frac{1}{B}\sum_{b = 1}^B F^{(b)}_n(x)$ and
$\mathbb{E}_B^{*}[\hat\theta^{(b)}_n]  =  \frac{1}{B}\sum_{b = 1}^B\hat\theta^{(b)}_n$.
\end{enumerate}
The p-value of the block bootstrap KS test can be approximated
as $p = \sum_{b=1}^B \mathbf{1}\{T^{(b)}_n > T_n\} / B$.

@@

Several promising directions for further research emerge from the findings of
this study. Firstly, there are many other variants of bootstrap besides
circular block bootstrap which
can be used for serially dependent data such as the
stationary bootstrap \citep{politis1994stationary} and the dependent wild 
bootstrap \citep{shao2010dependent}. It would be interesting to investigate how
these methods perform using our correction. Secondly, analogous 
methodologies for other goodness-of-fit tests
based on empirical distributions, such as the Anderson--Darling and Cramér--von
Mises tests---which have been shown to outperform the KS test in certain scenarios
\citep{stephens2017tests}---could be explored. The application of block
bootstrapping bias correction to these tests could be particularly beneficial in
environmental studies, such as extreme value modeling, where observations are
often serially dependent. Thirdly, an extension of the proposed NPBB
one-sample KS test to a two-sample setting would be
intriguing. This adaptation would test whether two samples originate from the
same distribution without requiring specification of serial dependencies in
either sample, maintaining the advantageous feature of the original method.
Finally, a rigorous theoretical justification of the NPBB method for
stationary series is of significant academic interest. Establishing
precise conditions for the asymptotic behavior of the KS statistic and
the consistency of the bootstrap procedure would offer valuable
practical insights into the scenarios where the NPBB approach is
expected to perform optimally.

### 2025-02-18 manuscript/reply.tex
politis2004blksize

Thank you. This is an interesting direction to investigate. We have added few
sentences concerning potential future investigation into evaluating a bias 
correction for these variants in the second paragraph of the Discussion 
section.

>>>>

Thank you. This is an interesting direction to investigate. We have added a few
sentences concerning potential future investigation into evaluating a bias 
correction for these variants in the second paragraph of the Discussion 
section.
